Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-02-25 draft.tex
Incorporate more of Ulrik's comments

Our initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare decay FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, note that the method we
describe in the following is general, and it applies to arbitrary decay
or scattering processes. We find a previous work \cite{Dighe:1998vk}
that advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays, but also mentions the
applicability to semileptonic decays. Our aim is to improve upon this previous
work by studying the uncertainties that are introduced by mismodelling
of the angular distribution, and by working out a recipe to determine
and unfold detector effects. Moreover, we show
that the method of moments has two major advantages over the usual
approach based on likelihood fits:
\begin{enumerate}
    \item \del{Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF, see e.g. \cite{Egede:1142152}.
        The method of moments has no such convergence problems.
    }
    \item \add{Likelihood fits of problems, for which the underlying
        physical model is only partially known, may be delicate. Such cases can lead to
        overestimating the number of physical parameters, and consequently
        inhibit the convergence of the fits. (As an example, this type
        of problem occured in toy studies of the decay $B\to K^*(\to K\pi)\ell^+\ell^-$
        as reported in \cite{Egede:1142152}. It was subsequently solved
        when a missing symmetry relation between the angular observables
        was found and applied, thereby reducing the number of fit parameters.)
    }
    \item The results of likelihood fits suffer from systematic uncertainties
        and systematic bias \del{(see e.g. \cite{Egede:1142152})}, especially in but not limited
        to cases in which the PDFs do not faithfully model the underlying physics. We will
        show that the method of moments does not suffer a very particular type of
        mismodelling that arises from introducing a cutoff in a partial wave expansion.
    \item Using the method of moments, the joint probability distribution of the
        angular observables converges towards a multivariate Gaussian distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

@@

Compared to the usual maximum-likelihood approach, we find for the method of moments:
\begin{enumerate}
  \item The angular observable ${S_i}$ can be determined
  independently of any other observable ${S_j}$. It is therefore
  much more robust to physics assumptions needed to define the full
  likelihood. In particular, this means one does not have to be
  specific regarding the form of new-physics contributions; in fact,
  one does not even need to be able to explicitly formulate the
  likelihood at all.
\item It is superior for a small number of samples $N$. Likelihood
  fits tend to be numerically unstable if lots of parameters need to
  be estimated from sparse data. This is more severe if the mode of
  the likelihood is near the boundary of the physically allowed
  region.\fred{Is there some LHCb note we could cite?}\danny{Ulrik suggests citing some
  statistical paper}
  For some of these decays of interest, there are only $\order{10}$--$\order{100}$ events
  per bin on record.\danny{What to cite? 1403.8004 only states total signal yield!}
\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator and vice versa~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. It is therefore highly questionable \add{to outright
  prefer the method of moments over likelihood fits, in order} to reduce
  only the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties (10--15)\%
  smaller than those from the method of moments.
\item The approximate multivariate Gaussian distribution of $\vecest{S}$
    allows easier and more \add{precise}\del{correct} transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
Descotes-Genon:2013wba,Beaujean:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  For experimentalists, communicating such a result has
  proved to be challenging due to technical reasons (data formats,
  size etc.) and the fear that theorists misinterpret the many
  nuisance parameters (from control channels, detector calibrations
  etc.) that are also part of the likelihood.
  \danny{Ulrik: Seems an odd comment to put in a paper.}
  This leads to the
  undesirable situation that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem as well as correlations altogether.

@@

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokstarll} for details
on the angular distribution.) However, at the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $j_z = -1,0,+1$.
It is then \add{convenient}\del{usual} to compute the angular distribution explicitly in terms of $\theta_1$ and $\theta_3$, but leave
the angular observables dependent on the remaining helicity angle $\theta_2$.
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{l}^{(|m|)}$
with respect to the remaining angle $\theta_2$. \add{Here, the angular momentum-like indices $l$ and $m$ follows from the usual
rules for addition of the angular momenta $j$ and $\tilde{j}$ of the partial wave expansion of the underlying
amplitude and its complex conjugate: $j - \tilde{j} \leq l \leq j + \tilde{j}$, and $m = j_z + \tilde{j}_z$.}
Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{l,|m|}} S_{k,l}(\vec{\nu}) p_{l}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{l,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokstarll} and \cite{Lee:1992ih})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,,\\
        1\,, & k = 4,5,7,8\,,\\
        2\,, & k = 3,9
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. \del{To wit}\add{That is to say},
ignoring further (orthogonal) terms in the PDF cannot affect $\vec{S}$ nor its variance.\\

@@

We wish to emphasize, however, that detector acceptance effects systematically affect the expansion for any basis
of angular functions, including the one suggested in this section.
\add{Nevertheless, the expansion in terms of Legendre polynomials provides means to cope with these effects,
as we discuss in the following subsection.}
\del{The means to cope with this problem are discussed
in the following subsection.}

@@

The determination of a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance effects and could thus
write down the full forward model from which the measured events
arise. In practice, this would require much more CPU time than
available, and one is therefore forced to simplify the model. The
standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$.  In general, the distribution of the detected
events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \frac{1}{R} \int P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}}\,.
\end{equation}
Here $P_\text{t}$ is the probability distribution of the true events,
the normalization constant $R$ is given by
\begin{equation}
    R \equiv \iint P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}} \rmdx{x_\text{d}}\,,
\end{equation}
and the kernel $E(x_\text{d}| x_\text{t})$ parametrizes acceptance and
resolution effects. The kernel is usually decomposed as
\begin{equation}
  \Ekernel = \eps(\xt) P(\xd|\xt),
\end{equation}
where the PDF $P(\xd|\xt)$ models the resolution effects and the
unnormalized density $\eps(\xt)$ is the \emph{detector acceptance}
function. Perfect resolution corresponds to
\begin{equation}
  \label{eq:def-R}
  P(\xd|\xt) = \delta(\xd - \xt) \,.
\end{equation}

In what follows, we propose a systematic method to unfold all effects
of $\Ekernel$ through MC simulations and the method of moments.
\add{In this process, we use that fact} that $\Ekernel$ is independent of the angular
observables, and that it is possible to expand $\Ekernel$ --- at
least formally --- in Legendre polymials.  For illustration, we
proceed with the explicit example of a uniangular\footnote{%
  The generalization of this section to multiangular PDFs is
  straighforward. It can be achieved by promoting $x$ to a vector,
  promoting the Legendre polynomials to products of independent
  polynomials or spherical harmonics, and promoting the indices
  $i,j,k,m$ to multi-indices.  } PDF with $x = \cos\theta$. Let us
define the PDF in terms of the Legendre polymials (i.e., $f_k(x)
\equiv P_k(x)$) and angular observables $\vec{S}$ as
\begin{equation}
    P\subt(\xt) \equiv P\subt(\xt | \vec{S}) = \sum_k S_k f_k(\xt),
  \end{equation}
where $k = 0, 1, \dots$ denotes an angular-momentum-like index associated with the observables.
Normalization of $P\subt$ is equivalent to choosing $S_0 = 1/2$. Requiring $P\subt(\xt) \ge 0 \,\forall \xt$
implies $|S_k| \leq 1/2$ for $k > 0$. More stringent relations between the
$S_k$ might hold, but are beyond the scope of this work. For later use, we define
\begin{equation}
  \label{eq:def-Sk}
  S^{(m)}_k \equiv
  \begin{cases}
    1/2 \delta_{k,0}, & m=0\\
    1/2 (\delta_{k,0} + \delta_{k,m}), & m>0 \,,
  \end{cases}
\end{equation}
and note that
\begin{equation}
    P^{(m)}(\xt) \equiv P(\xt | \vec{S}^{(m)})
\end{equation}
is a valid PDF. The dual basis of angular functions follows then from the normalization
and orthogonality of the Legendre polynomials, and one therefore has
\begin{equation}
    \tilde{f}_k(x) = \frac{2 k + 1}{2} f_k(x),\qquad \int_{-1}^{+1} \tilde{f}_k(x) f_l(x) \rmdx{x} = \delta_{k,l}\,.
\end{equation}
We now define the \emph{simulated raw} moments $\vec{Q}^{(m)}$ as
\begin{equation}
    Q_i^{(m)} \equiv \iint \tilde{f}_i(\xd) P^{(m)}(\xt) \Ekernel \rmdx{\xt} \rmdx{\xd},
\end{equation}
which are instrumental to our recipe. Monte Carlo estimators of these moments can be constructed from specifically crafted detector events $x_\text{d}^{(n,m)}$, $n = 1, \dots, N_\text{d}^{(m)}$, where
\begin{equation}
    x_\text{d}^{(n,m)} \sim P_E^{(m)}(\xd) \equiv \frac{1}{R^{(m)}} \int P^{(m)}(\xt) \Ekernel \rmdx{\xt}\,.
\end{equation}
In words, for each $m$ it is required to generate events from
a toy physical distribution $P^{(m)}$ in which $S_0 = S_m=1/2$ and all other
observables are set to zero, and to propagate these events through
a detector simulation. The normalization $R^{(m)}$ is chosen such
that $\int P_E^{(m)}(\xd) \rmdx{\xd} = 1$.  We emphasize that
$R^{(m)}$ can be estimated as $\est{R}^{(m)} = N_\text{d}^{(m)} /
N_\text{t}$, where $N_\text{t}$ corresponds to the number of
simulated true events.  The estimators then read
\begin{equation}
    \est{Q}_i^{(m)} \equiv \est{R}^{(m)} \frac{1}{N_\text{d}} \sum_n^{N_\text{d}} \tilde{f}_i(x_\text{d}^{(n,m)})
\end{equation}

@@

In order to finally extract the angular observables from data, we use the \emph{measured raw moments}. Their MC estimator
$\vecest{Q}$ --- based on the \emph{detected events} $x^{(n)}$, $n=1,\dots,N$ --- reads
\begin{equation}
    \est{Q}_i \equiv R \frac{1}{N} \sum_n^N \tilde{f}_i(x^{(n)})\,.
\end{equation}
We then obtain MC estimators of the angular observables via
\begin{equation}
    \label{eq:estS}
    \vecest{S}_i \equiv \left[\est{M}^{-1}\right] \vecest{Q}\,.
\end{equation}
The precise value of $R$ (see \refeq{def-R}) depends on the a priori
unknown angular observables $\vec S$.  However, ratios and
other suitable combinations of the estimators $\vecest{S}$
are independent of $R$. Conversely, $R$ can be
extracted from the self-consistency condition\footnote{This condition
  is process dependent; see \refeq{recipe:lambdabtolambdall} and
  \refeq{recipe:btokstarll}.}  $\hat{S}_0 = \frac{1}{2}$. Since $R$ is
relevant in the determination of the branching ratio, it is therefore
useful to carry out that determination and the extraction of the
angular observables in a combined analysis.\\

We note that $\vecest{S}$ as determined from \refeq{estS} does not fully correspond to $\vec{S}$
for arbitrary $\eps$. Taking again the decay $B\to K\ell^+\ell^-$
as an example, \del{we have} \add{there are exactly three physical angular observables:} $\dim S = 3$.
Assuming an expansion of the detector acceptance $\eps$
in terms of Legendre polynomial up to order $L$, we have to calculate the raw moments up
to $\dim \est{Q} = \dim \est{S} = \dim S + L$. The MC estimators of the corrected angular observables
then take the following structure:
\begin{equation}
    \vecest{S} = ( \underbrace{\est{S}_0, \dots, \est{S}_{\dim S}}_\text{physical}, \underbrace{\est{S}_{\dim S + 1}, \dots, \est{S}_{\dim Q}}_\text{``superfluous''})\,.
\end{equation}
The method is consistent as long as the MC estimators for the ``superfluous'' observables with indices $\dim S + 1$ to $\dim Q$
are compatible with zero. \footnote{This holds only for the physical model of uniangular decay distributions. For some decays, these superfluous observables
correspond to higher partial waves that have been suppressed in the physical model; see \refapp{btokstarll} for such a case.}
Since $L$ depends on the setup of the particle detector under consideration, it remains for
the experimental collaborations to determine the maximum order of such an expansion.
\add{We propose to do this by \dots}\\

>>>>

Our initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare decay FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, note that the method we
describe in the following is general, and it applies to arbitrary decay
or scattering processes. We find a previous work \cite{Dighe:1998vk}
that advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays, but also mentions the
applicability to semileptonic decays. Our aim is to improve upon this previous
work by studying the uncertainties that are introduced by mismodelling
of the angular distribution, and by working out a recipe to determine
and unfold detector effects. Moreover, we show
that the method of moments has \add{several} major advantages over the usual
approach based on likelihood fits:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF. \add{As an example,
        see the LHCb analysis of \dots}.
        The method of moments does not require reparametrization and approximations.

    \item \add{Likelihood fits of problems, for which the underlying
        physical model is only partially known, may be delicate. Such cases can lead to
        overestimating the number of physical parameters, and consequently
        inhibit the convergence of the fits. (As an example, this type
        of problem occured in toy studies of the decay $B\to K^*(\to K\pi)\ell^+\ell^-$
        as reported in \cite{Egede:1142152}. It was subsequently solved
        when a missing symmetry relation between the angular observables
        was found and applied, thereby reducing the number of fit parameters.)
    }
    \item Mismodeling the underlying physics model can result in systematic bias
        in likelihood fits. We will
        show that the method of moments is insensitive to a certain type of mismodeling;
        i.e. introducing a cutoff in a partial-wave expansion of the signal PDF.

    \item Using the method of moments, the joint probability distribution of the
        angular observables rapidly converges towards a multivariate Gaussian distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

@@

Compared to the usual maximum-likelihood approach, we find for the method of moments:
\begin{enumerate}
  \item The angular observable ${S_i}$ can be determined
  independently of any other observable ${S_j}$. It is therefore
  much more robust to physics assumptions needed to define the full
  likelihood. In particular, this means one does not have to be
  specific regarding the form of new-physics contributions; in fact,
  one does not even need to be able to explicitly formulate the
  likelihood at all.
\item It is superior for a small number of samples $N$. Likelihood
  fits tend to be numerically unstable if lots of parameters need to
  be estimated from sparse data. This is more severe if the mode of
  the likelihood is near the boundary of the physically allowed
  region.\fred{Is there some LHCb note we could cite?}\danny{Ulrik suggests citing some
  statistical paper}
  For some of these decays of interest, there are only $\order{10}$--$\order{100}$ events
  per bin on record.\danny{What to cite? 1403.8004 only states total signal yield!}
\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator and vice versa~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. \add{One should therefore not prefer the method of moments
  over likelihood fits just because the former reduces}
  the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties \add{(10--15)\%}
  smaller than those from the method of moments.
\item The approximate multivariate Gaussian distribution of $\vecest{S}$
    allows easier and more \add{precise} transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
Descotes-Genon:2013wba,Beaujean:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  Communicating such a result has
  proved to be challenging due to technical reasons (such as data formats,
  size etc.) \add{as well as political reasons; i.e., the publication of
  the likelihood also publishes collaboration-internal data.}
  This leads to the
  undesirable situation that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem as well as correlations altogether.

@@

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokstarll} for details
on the angular distribution.) However, at the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $j_z = -1,0,+1$.
It is then \add{convenient}\del{usual} to compute the angular distribution explicitly in terms of $\theta_1$ and $\theta_3$, but leave
the angular observables dependent on the remaining helicity angle $\theta_2$.
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{l}^{(|m|)}$
with respect to the remaining angle $\theta_2$. \add{Here, the angular momentum-like indices $l$ and $m$ follows from the usual
rules for addition of the angular momenta $j$ and $\tilde{j}$ of the partial wave expansion of the underlying
amplitude and its complex conjugate: $j - \tilde{j} \leq l \leq j + \tilde{j}$, and $m = j_z + \tilde{j}_z$.}
Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{l,|m|}} S_{k,l}(\vec{\nu}) p_{l}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{l,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokstarll} and \cite{Lee:1992ih})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,,\\
        1\,, & k = 4,5,7,8\,,\\
        2\,, & k = 3,9
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. \add{That is to say},
ignoring further (orthogonal) terms in the PDF cannot affect $\vecest{S}$ nor its variance.\\

@@

We wish to emphasize, however, that detector acceptance effects systematically affect the expansion for any basis
of angular functions, including the one suggested in this section.
Nevertheless, the expansion in terms of Legendre polynomials as suggested above provides means to cope with these effects,
as we discuss in the following subsection.

@@

The determination of a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance effects and could thus
write down the full forward model from which the measured events
arise. In practice, that is not feasible, and one is
forced to simplify the model.
The standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$.  In general, the distribution of the detected
events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \frac{1}{R} \int P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}}\,.
\end{equation}
Here $P_\text{t}$ is the probability distribution of the true events,
the normalization constant $R$ is given by
\begin{equation}
  \label{eq:def-R}
    R \equiv \iint P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}} \rmdx{x_\text{d}}\,,
\end{equation}
and the kernel $E(x_\text{d}| x_\text{t})$ parametrizes acceptance and
resolution effects. The kernel is usually decomposed as
\begin{equation}
  \Ekernel = \eps(\xt) P(\xd|\xt),
\end{equation}
where the PDF $P(\xd|\xt)$ models the resolution effects and the
unnormalized density $\eps(\xt)$ is the \emph{detector acceptance}
function. Perfect resolution corresponds to
\begin{equation}
  P(\xd|\xt) = \delta(\xd - \xt) \,.
\end{equation}

In what follows, we propose a systematic method to unfold all effects
of $\Ekernel$ through MC simulations and the method of moments,
using that $\Ekernel$ can be expanded --- at
least formally --- in Legendre polymials.  For illustration, we
proceed with the explicit example of a uniangular\footnote{%
  The generalization of this section to multiangular PDFs is
  straighforward. It can be achieved by promoting $x$ to a vector,
  promoting the Legendre polynomials to products of independent
  polynomials or spherical harmonics, and promoting the indices
  $i,j,k,m$ to multi-indices.  } PDF with $x = \cos\theta$. Let us
define the PDF in terms of the Legendre polymials (i.e., $f_k(x)
\equiv P_k(x)$) and angular observables $\vec{S}$ as
\begin{equation}
    P\subt(\xt) \equiv P\subt(\xt | \vec{S}) = \sum_k S_k f_k(\xt),
  \end{equation}
where $k = 0, 1, \dots$ denotes an angular-momentum-like index associated with the observables.
Normalization of $P\subt$ is equivalent to choosing $S_0 = 1/2$. Requiring $P\subt(\xt) \ge 0 \,\forall \xt$
implies $|S_k| \leq 1/2$ for $k > 0$. More stringent relations between the
$S_k$ might hold, but are of no concern here. For later use, we define
\begin{equation}
  \label{eq:def-Sk}
  S^{(m)}_k \equiv
  \begin{cases}
    1/2 \delta_{k,0}, & m=0\\
    1/2 (\delta_{k,0} + \delta_{k,m}), & m>0 \,,
  \end{cases}
\end{equation}
and note that
\begin{equation}
    P^{(m)}(\xt) \equiv P(\xt | \vec{S}^{(m)})
\end{equation}
is a valid PDF. The dual basis of angular functions follows then from the normalization
and orthogonality of the Legendre polynomials, and one therefore has
\begin{equation}
    \tilde{f}_k(x) = \frac{2 k + 1}{2} f_k(x),\qquad \int_{-1}^{+1} \tilde{f}_k(x) f_l(x) \rmdx{x} = \delta_{k,l}\,.
\end{equation}
We now define the \emph{simulated raw} moments $\vec{Q}^{(m)}$ as
\begin{equation}
    \label{eq:analytic-raw-moments}
    Q_i^{(m)} \equiv \iint \tilde{f}_i(\xd) P^{(m)}(\xt) \Ekernel \rmdx{\xt} \rmdx{\xd},
\end{equation}
which are instrumental to our recipe. Monte Carlo estimators of these moments can be constructed from specifically crafted detector events $x_\text{d}^{(n,m)}$, $n = 1, \dots, N_\text{d}^{(m)}$, where
\begin{equation}
    x_\text{d}^{(n,m)} \sim P_E^{(m)}(\xd) \equiv \frac{1}{R^{(m)}} \int P^{(m)}(\xt) \Ekernel \rmdx{\xt}\,.
\end{equation}
In words, for each $m$ it is required to generate events from
a toy physical distribution $P^{(m)}$ in which $S_0 = S_m=1/2$ and all other
observables are set to zero, and to propagate these events through
a detector simulation. The normalization $R^{(m)}$ is chosen such
that $\int P_E^{(m)}(\xd) \rmdx{\xd} = 1$.  We emphasize that
$R^{(m)}$ can be estimated as $\est{R}^{(m)} = N_\text{d}^{(m)} /
N_\text{t}$, where $N_\text{t}$ corresponds to the number of
simulated true events.  The estimators then read
\begin{equation}
    \est{Q}_i^{(m)} \equiv \est{R}^{(m)} \frac{1}{N_\text{d}} \sum_n^{N_\text{d}} \tilde{f}_i(x_\text{d}^{(n,m)})
\end{equation}

@@

In order to finally extract the angular observables from data, we use the \emph{measured raw moments}. Their MC estimator
$\vecest{Q}$ --- based on the \emph{detected events} $x^{(n)}$, $n=1,\dots,N$ --- reads
\begin{equation}
    \est{Q}_i \equiv R \frac{1}{N} \sum_n^N \tilde{f}_i(x^{(n)})\,.
\end{equation}
We then obtain MC estimators of the angular observables via
\begin{equation}
    \label{eq:estS}
    \vecest{S} \equiv \left[\est{M}^{-1}\right] \vecest{Q}\,.
\end{equation}

\add{Apparently we now face a circular dependence. On the one hand,
the estimators $\vecest{Q}$, and thus also the estimator $\vecest{S}$
are proportional to $R$, the ratio of detected events over occuring events.
On the other hand, $R$ depends by construction (see \refeq{def-R})
on $P_\text{t}(x_t)$, and thus on the true value of the angular observables $\vec{S}$.
We solve this problem by observing that the MC estimators $\vecest{S}$ need to fulfill
a self-consistency condition, which reads
\begin{equation}
    \est{S}_0 = \frac{1}{2}\,,\qquad\text{for a uniangular distribution}.
\end{equation}
(For the process-dependent conditions in the multiangular case see \refeq{recipe:lambdabtolambdall} and
\refeq{recipe:btokstarll}.)
This self-consistency condition is tightly related to the determination of
the branching ratio of the underlying decay, and we therefore suggest to
carry out a combined analysis for the determination of the branching ratio
and the extraction of the angular observables.\\
Moreover, in the applications to $B$ decays only ratios and similar $R$-independent combinations
of the angular observables are of interest.
}\\

\del{The precise value of $R$ (see \refeq{def-R}) depends on the a priori
unknown angular observables $\vec S$.  However, within ratios and
other suitable combinations of the estimators $\vecest{S}$, the quantity
$R$ cancels.
Conversely, $R$ can be
extracted from the self-consistency condition\footnote{This condition
  is process dependent; see \refeq{recipe:lambdabtolambdall} and
\refeq{recipe:btokstarll}.}  $\hat{S}_0 = \frac{1}{2}$. Since $R$ is
relevant in the determination of the branching ratio, it is therefore
useful to carry out that determination and the extraction of the
angular observables in a combined analysis.}\\

We note that $\vecest{S}$ as determined from \refeq{estS} does not fully correspond to $\vec{S}$
for arbitrary $\eps$. Taking again the decay $B\to K\ell^+\ell^-$
as an example, \add{there are exactly three physical angular observables:} $\dim S = 3$.
Assuming an expansion of the detector acceptance $\eps$
in terms of Legendre polynomial up to a given order $L$, we have to calculate the raw moments up
to $\dim \est{Q} = \dim \est{S} = \dim S + L$. The MC estimators of the corrected angular observables
then take the following structure:
\begin{equation}
    \vecest{S} = ( \underbrace{\est{S}_0, \dots, \est{S}_{\dim S}}_\text{physical}, \underbrace{\est{S}_{\dim S + 1}, \dots, \est{S}_{\dim Q}}_\text{``superfluous''})\,.
\end{equation}
The method is consistent as long as the MC estimators for the ``superfluous'' observables with indices $\dim S + 1$ to $\dim Q$
are compatible with zero. \footnote{This holds only for the physical model of uniangular decay distributions. If the physical
    model involves a partial-wave expansion with cutoff, these superfluous observables
correspond to higher partial waves that have been suppressed in the physical model; see \refapp{btokstarll} for such a case.}
Since $L$ depends on the setup of the particle detector under consideration, it remains for
the experimental collaborations to determine the maximum order of such an expansion.
\add{We suggest to carry out a dedicated study to determine a suitable choice for $L$.
Since such a work requires in-depth understanding of the detector simulation, it remains
beyond the scope of our possibilities.}\\
